Screening of airborne radioxenon measurements for Nuclear-Test-Ban Treaty verification

### 2022-10-24 paper/paper.tex
fix grammar anomaly detection

>>>>

\section{Anomaly detection}

One of the main challenges of applying traditional anomaly detection methods on the time-series data of an observation station, is accommodating for weather. Winds will change the course of an emission, causing different observation stations to pick up their signature. Therefore treating stations individually to try to learn their normal behavior would not lead to fruitful results. At the same time, it is also difficult to train more complex models that handle data collectively from all stations, since there is not enough historic data, and only a few observation stations are actively collecting data throughout the entire year.\\

For these reasons, working with the emission sources directly is more appealing. Correcting the emission data in the previous chapters allows us to obtain a better ground truth for the variation of the emissions over the course of the year. The resulting individual time-series data should show a few characteristics that arise from the nature of these sources:
\begin{itemize}
    \item For power plants, a long-term baseline variation corresponding to yearly electricity demand, which should vary consistently given that our readings are all in the northern hemisphere.
    \item For medical isotope production facilities, a short term periodic trend corresponding to the working days of the week.  
\end{itemize}

After adjusting for the yearly trend, the eventual solution should include two components. The first which will segment the time-series into pieces with distinct behavior or features, for example using shapelets \cite{10.1145/2623330.2623613}\cite{10.1007/978-3-030-39098-3_7} or change point detection algorithms \cite{TRUONG2020107299}. The second part is to train anomaly detection models or classifiers, that distinguish between segments of normal and abnormal behavior. The state of the art methods use 1D convolution neural networks \cite{KIRANYAZ2021107398} or bayesian probabilistic models \cite{https://doi.org/10.48550/arxiv.1902.08627}. Finally, synthetic emission data should be simulated in order to evaluate these methods, since data for nuclear tests is rare if not inexistent.
